---
title: "Re-evaluating Growth in the Time of Debt"
subtitle: "[DSLC stages]: Data cleaning and pre-processing"
format: 
  html:
    css: theme.css
    toc: true
    toc-location: right
    number-depth: 4
    theme: cerulean
    df-print: kable
execute:
  echo: true
editor: source
number-sections: true
embed-resources: true
editor_options: 
  chunk_output_type: console
---


## Domain problem formulation
The goal is to try and reproduce the findings of Reinhart and Rogoff’s study (introduced in Chapter 3) using historical public debt data as well as gross 
domestic product (GDP) growth data. 

## Data source overview
historical public debt data (debt.xls) from the International Monetary Fund (IMF),  
Gross domestic product (GDP) growth data (growth.csv).


## Step 1: Review background information {#sec-bg-info}
``` {r}

```

### Data dictionary
``` {r}
```

## Step 2: Loading in the data
``` {r}
#install.packages("readxl")
#library(readxl)
library(tidyverse)
library(dplyr)

#Data first was fixed in Google sheets, and now I am loading corrected data
growth_original <- read_csv('/Users/romansemenko/Documents/FALL 2025/Practical data analysis/HW5/growth_original.csv')
head(growth_original)
debt_original <- read_csv('/Users/romansemenko/Documents/FALL 2025/Practical data analysis/HW5/debt_original.csv')

```

## Step 3: Examine the data and create action items
"No data" values should be "NA" 
Data should be converted to long format
###Debt:
remove data for "debt" before 1960
first columns should be renamed to "country"
convert debt % rate, year from char to numeric
###Growth 
remove non-existing countries
renaming countries that have different name in the debt table
convert growth % rate, year char to numeric
### Finding invalid values
``` {r}
#Debt
source('prepare_debt_data.R')
debt_clean <- prepare_debt_data(debt_original)
#Growth 
source("prepare_growth_data.R")
growth_clean <- prepare_growth_data(growth_original)
```

### Assessing Column names
Done above


### Assessing variable type
done before 

### Assessing data completeness


``` {r}
#For the DEBT data
dim(debt_clean)
#note it's 194 (countries) times 56 years 

debt_clean |> 
  # keep only numeric variables
  select(where(is.numeric)) |> 
  # for each column, compute a data frame with the min, mean, and max.
  map_df(function(.col) { data.frame(min = min(.col, na.rm = TRUE),
                                     mean = mean(.col, na.rm = TRUE),
                                     max = max(.col, na.rm = TRUE)) },
         .id = "variable")


debt_clean %>%
  sample_n(10)



#For the GROWTH data
dim(growth_clean)


growth_clean |> 
  # keep only numeric variables
  select(where(is.numeric)) |> 
  # for each column, compute a data frame with the min, mean, and max.
  map_df(function(.col) { data.frame(min = min(.col, na.rm = TRUE),
                                     mean = mean(.col, na.rm = TRUE),
                                     max = max(.col, na.rm = TRUE)) },
         .id = "variable")


growth_clean %>%
  sample_n(10)
```


###Comparing Australia as on gradescope
```{r}
debt_clean %>% filter(country == 'Australia')

growth_clean %>% filter(country == 'Australia') %>% 
  select(country, year, growth_as_percent)
```


## Step 4: Clean and pre-process the data
done in steps before
Now merging two tables
```{r}
source("merge_two_tables.R")
result <- merge_two_tables(growth_clean,debt_clean)

#Check the result for Australia
result %>% filter(country == 'Australia') %>%
  select(country,year,growth_as_percent,debt_as_percent_of_gdp)
```
